### 2017-11-16 proposal.tex
proposal: add outline

\documentclass{article}
\usepackage[tt=false]{libertine}
\usepackage{hyperref}
\usepackage{mathpartir}
\usepackage{amsthm}
\usepackage{fullpage}

\usepackage{sectsty}
\allsectionsfont{\sffamily}

\title{Implementing DNS in C0}
\author{Farzaneh Derakhshan, Klaas Pruiksma, Jonathan Sterling}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\newcommand\Lang[1]{{\sffamily\bfseries{#1}}}
\newcommand\LangCA{\Lang{CA}}
\newcommand\LangMA{\Lang{MA}}
\newcommand\EOf[4]{{#1}\vdash_{#2}{#3}:{#4}}
\newcommand\MOf[4]{{#1}\vdash_{#2}{#3}\div{#4}}
\newcommand\IsProc[3]{{#1}\vdash_{#2}{#3}\ \mathsf{proc}}
\newcommand\IsAction[3]{{#1}\vdash_{#2}{#3}\ \mathsf{action}}

\begin{document}

\maketitle

\begin{abstract}
  Please write an abstract.
\end{abstract}

\section{Background and Related Work}

\subsection{What is DNS?}

Jon

\subsection{C0 and Concurrent C0}

Farzaneh.

\subsection{Related Work}
Jon.

\section{Our Project}

\subsection{Network interface}

Klaas.

\subsection{Marshalling and unmarshalling}

Jon

\section{Evaluation Plan}

Klaas.

\bibliographystyle{abbrv}
\bibliography{bibtex-references/refs}

\end{document}